lfl: An \R{} Package for Linguistic Fuzzy Logic

### 2021-06-16 article.tex
Updates from Overleaf

The aim of this paper is to present particular package for the \R{} statistical environment \citep{R2020, Rintro} named {\bf lfl} that enables the use of the linguistic fuzzy logic in data processing applications. The package provides implemented tools for using the results of original work of \cite{Novak08, Novak:PbLD, DvoStep:PbLD2015, StepBurda:FRBE_FSS}, and others, and it provides executable routines that are not freely available anywhere else.

@@

\begin{table}[]
    \centering
    \begin{tabular}{l|rr}
        rule & support & confidence \\
        \hline
                                 {Sepal.Length=(7.18,7.9]} $\Rightarrow$ {Species=virginica} & 0.073 & 1.000 \\
                                {Petal.Length=(5.72,6.91]} $\Rightarrow$ {Species=virginica} & 0.107 & 1.000 \\
                                  {Petal.Width=(2.02,2.5]} $\Rightarrow$ {Species=virginica} & 0.153 & 1.000 \\
                                  {Petal.Width=(0.0976,0.58]} $\Rightarrow$ {Species=setosa} & 0.327 & 1.000 \\
                                  {Petal.Length=(0.994,2.18]} $\Rightarrow$ {Species=setosa} & 0.333 & 1.000 \\
        {Sepal.Length=(7.18,7.9] \& Petal.Length=(5.72,6.91]} $\Rightarrow$ {Species=virginica} & 0.073 & 1.000 \\
         {Petal.Length=(5.72,6.91] \& Petal.Width=(2.02,2.5]} $\Rightarrow$ {Species=virginica} & 0.060 & 1.000 \\
        {Sepal.Width=(2.96,3.44] \& Petal.Length=(5.72,6.91]} $\Rightarrow$ {Species=virginica} & 0.053 & 1.000 \\
    \end{tabular}
    \caption{First 8 rules with the greatest confidence as found by the \pkg{arules} package.}
    \label{tab:arules8rules}
\end{table}

As can be seen, both packages provide comparable  results. For \pkg{arules}, the numeric data have to be categorized into crisp intervals. The \lfl{} package allows to work with linguistic expressions modelled with fuzzy sets, which enables non-sharp borders between categories. The interpretability as well as the lower number of rules play in favor of \pkg{lfl} however, we do not dare to generalize too much from the latter observation as each case study could lead to a different result and the impact of the a proper categorization is essential. 
\Martin{pridal jsem posledni vetu, kouukni na ni. Jinak jsem to nove opet cele prosel a mistz opravil nebo upravil. Nevim, zda jeste hodlas menit ta pravidla do tabulky, jinak jsem ready submitovat.}

>>>>

The aim of this paper is to present particular package for the \R{} statistical environment \citep{R2020, Rintro} named \pkg{lfl} that enables the use of the linguistic fuzzy logic in data processing applications. The package provides implemented tools for using the results of original work of \cite{Novak08, Novak:PbLD, DvoStep:PbLD2015, StepBurda:FRBE_FSS}, and others, and it provides executable routines that are not freely available anywhere else.

@@

\begin{table}[t]
    \centering
    \begin{tabular}{l|rr}
        rule & support & confidence \\
        \hline
                                 {Sepal.Length=(7.18,7.9]} $\Rightarrow$ {Species=virginica} & 0.073 & 1.000 \\
                                {Petal.Length=(5.72,6.91]} $\Rightarrow$ {Species=virginica} & 0.107 & 1.000 \\
                                  {Petal.Width=(2.02,2.5]} $\Rightarrow$ {Species=virginica} & 0.153 & 1.000 \\
                                  {Petal.Width=(0.0976,0.58]} $\Rightarrow$ {Species=setosa} & 0.327 & 1.000 \\
                                  {Petal.Length=(0.994,2.18]} $\Rightarrow$ {Species=setosa} & 0.333 & 1.000 \\
        {Sepal.Length=(7.18,7.9] \& Petal.Length=(5.72,6.91]} $\Rightarrow$ {Species=virginica} & 0.073 & 1.000 \\
         {Petal.Length=(5.72,6.91] \& Petal.Width=(2.02,2.5]} $\Rightarrow$ {Species=virginica} & 0.060 & 1.000 \\
        {Sepal.Width=(2.96,3.44] \& Petal.Length=(5.72,6.91]} $\Rightarrow$ {Species=virginica} & 0.053 & 1.000 \\
    \end{tabular}
    \caption{First 8 rules with the greatest confidence as found by the \pkg{arules} package.}
    \label{tab:arules8rules}
\end{table}

As can be seen, both packages provide comparable  results. For \pkg{arules}, the numeric data have to be categorized into crisp intervals. The \lfl{} package allows to work with linguistic expressions modelled with fuzzy sets, which enables non-sharp borders between categories. The interpretability as well as the lower number of rules play in favor of \pkg{lfl} however, we do not dare to generalize too much from the latter observation as each case study could lead to a different result and the impact of the a proper categorization is essential.
